---
title: "Final Project"
author: "Qiyun Huang"
date: "2021/12/2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

#Loading pacakge
```{r }
pacman::p_load(tidyverse, ROCR, odds.n.ends, blorr, lmtest, car, summarytools, table1)
```


#importing data
```{r}
#yrbss2019 is the name of data frame I created in R, and yrbs2019.sav is the file name of SPSS dataset I want to import.
#use.value.label=TRUE to convert variables with value labels in SPSS into R factors.
#to.data.frame=TRUE to make as data frame.
yrbss2019 <- read.csv("https://raw.githubusercontent.com/baymaxxx97/gitdemo/main/yrbs2019.csv")
```


#Creating a smaller dataframe which only include the my data of interests
```{r}
yrbss2019_project <- yrbss2019[,c(1:2, 78,88,95,227,228,233)]
```

#Data management for variables used in this project (Q1,Q2,Q3,Q4,Q6,Q78,Q88,Q95, RACEETH)
```{r}
##provide names for variables
names(yrbss2019_project)<-c("age", "gender", "PA", "sleep_duration", "mus_stren", "obese","overweight", "race")


## Recode recode sleep duration into a binary variable – Sleep Deprivation
freq(yrbss2019_project$sleep_duration) ###cheking the varible frequency
yrbss2019_project$sleep_deprive <- ifelse(yrbss2019_project$sleep_duration == "7 hours" |
                                          yrbss2019_project$sleep_duration == "6 hours" |
                                          yrbss2019_project$sleep_duration == "5 hours" |
                                          yrbss2019_project$sleep_duration == "4 or less hours" , 1, 
                                   ifelse(yrbss2019_project$sleep_duration == "8 hours" |
                                          yrbss2019_project$sleep_duration == "9 hours" |
                                          yrbss2019_project$sleep_duration == "10 or more hours", 0, NA))
#lable binary variable sleep_deprive
yrbss2019_project$sleep_deprive <- ordered(yrbss2019_project$sleep_deprive,
levels = c(0, 1),
labels = c("Not Depreived", "Depreived Sleep"))

freq(yrbss2019_project$sleep_deprive) ###cheking the varible frequency

##check to make sure re-classification worked
table(yrbss2019_project$sleep_deprive, yrbss2019_project$sleep_duration)


## Age (Q1), Sex (Q2), and BMI are used to determine Overweight* (QNOWT), and Obese* (QNOBESE). The student is considered overweight, and QNOWT is set to “1”, when BMI percentile is at or above the 85th percentile and below the 95th        percentile for BMI by age and sex. The student is considered obese, and QNOBESE is set to “1”, when BMI percentile is at or above the 95th percentile for BMI by age and sex. If BMI is missing, then QNOWT and QNOBESE are set to missing. QNOWT and QNOBESE are mutually exclusive. The BMI percentile for age and sex (BMIPCT), QNOWT, and QNOBESE variables are included in the edited data set.

## Recode 1 into "Overweight/Obese" 0 into "not overweitht/not obese"
yrbss2019_project$obese_new <- ifelse(yrbss2019_project$obese == 1, 1,0)
yrbss2019_project$overweight_new <- ifelse(yrbss2019_project$overweight == 1, 1, 0)                         

#lable binary variable obese_new
yrbss2019_project$obese_new <- ordered(yrbss2019_project$obese_new,
levels = c(0, 1),
labels = c("Not Obese", "Obese"))

#lable binary variable overweight_new
yrbss2019_project$overweight_new <- ordered(yrbss2019_project$overweight_new,
levels = c(0, 1),
labels = c("Not Overweight", "Overweight"))

freq(yrbss2019_project$obese_new)
freq(yrbss2019_project$overweight_new)

##Checking frequency and class for variable PA and mus_stren
freq(yrbss2019_project$PA)
class(yrbss2019_project$PA)
freq(yrbss2019_project$mus_stren)
class(yrbss2019_project$mus_stren)


# exclude missing data
yrbss2019_project[yrbss2019_project == "Missing"] <- NA
yrbs_omit<-na.omit(yrbss2019_project)
```

#Creating Table 1
```{r}
#Addinf lables
label(yrbs_omit$age)<-"Age (years)"
label(yrbs_omit$gender)<-"Sex"
label(yrbs_omit$race)<-"Race"
label(yrbs_omit$PA)<-"Physical Activity"
label(yrbs_omit$mus_stren)<-"Muscle Strengthening"
label(yrbs_omit$obese_new)<-"Obese"
label(yrbs_omit$overweight_new)<-"Overweight"

table1(~age + gender + race+ PA + mus_stren + obese_new + overweight_new |sleep_deprive, overall="Total",  rowlabelhead="Variable", footnote='SD = standard deviation', yrbs_omit)
```




# Unajusted Logistic model for Physical activity
```{r}
#logistic model with bmi as a predictor
PA_Logit <- glm(sleep_deprive ~ PA, data=yrbs_omit, family="binomial")
summary(PA_Logit)

#calculate and print ORs and 95% CIs  
ORPA<-exp(cbind(OR = coef(PA_Logit), confint(PA_Logit))) #calculate ORs and 95% CIs
ORPA #print ORs and 95% CIs
```

“testing for multicollinearity are not necessary because we only have one predictor”

##Multivariate model with sleep deprivation as the dependent variable and Physical Activity as the independent variables
```{r}
#income and bmi cat logistic model
PALogit_multi <- glm(sleep_deprive ~ age + gender +race + PA, data=yrbs_omit, family="binomial")
summary(PALogit_multi)

#calculate and print ORs and 95% CIs  
ORmodel<-exp(cbind(OR = coef(PALogit_multi), confint(PALogit_multi))) #calculate ORs and 95% CIs
ORmodel #print ORs and 95% CIs

odds.n.ends(PALogit_multi)
```

## Look at assumptions of multicollinearity and influence

*Multicollinearity*
```{r}
#Variance Inflation Factors
vif(PALogit_multi)
```
#Cook's D plot
```{r}
plot(PALogit_multi, which=4, id.n=5, col="red") 
```

*Influence*
```{r}
#Cutoff: F(0.5, p, n-p), 4/n, 4/(n-p-1), n is the number of observations, p is the number of  coefficients
cutoff <- 0.0015 

#identify observations with a Cook's Distance greater than cutoff
obs_no <- as.data.frame(cooks.distance(PALogit_multi)) %>%
  mutate(obs_no=row_number()) %>%
  filter(`cooks.distance(PALogit_multi)` > cutoff)

#Let's exclude the values shown in the Cook's D plot, and see how the models compare.

PALogit_multi.modex <- update(PALogit_multi,subset=c(-obs_no$obs_no))
summary(PALogit_multi.modex)

#compare coefficients between models with and without influential observations, #caveat model number of observations is not the same
compareCoefs(PALogit_multi, PA_Logit) 
```


## Model fit of the full model
```{r}
#Various pseudo R squares, log likelihood, deviance, AIC, BIC
blr_model_fit_stats(PALogit_multi)

#Hosmer lemeshow goodness of fit test: a significant p value indicates a bad fit
blr_test_hosmer_lemeshow(PALogit_multi)
```
Based on the model results form PALogit_multi, we can conclude that after adjusting for age, gender, race and exclude the potential outliers are still statistically significant risk factors for sleep deprivation. 


#Unajusted Logistic model for Muscle Strength
```{r}
#income and bmi cat logistic model
MULogit <- glm(sleep_deprive ~ mus_stren, data=yrbs_omit, family="binomial")
summary(MULogit)

#calculate and print ORs and 95% CIs  
ORmodel<-exp(cbind(OR = coef(MULogit), confint(MULogit))) #calculate ORs and 95% CIs
ORmodel #print ORs and 95% CIs

odds.n.ends(MULogit)
```

##Multivariate model with sleep deprivation as the dependent variable and muscle strengthening the independent variables

```{r}
#income and bmi cat logistic model
MULogit_multi <- glm(sleep_deprive ~ age + gender +race + mus_stren, data=yrbs_omit, family="binomial")
summary(MULogit_multi)

#calculate and print ORs and 95% CIs  
ORmodel<-exp(cbind(OR = coef(MULogit_multi), confint(MULogit_multi))) #calculate ORs and 95% CIs
ORmodel #print ORs and 95% CIs

odds.n.ends(MULogit_multi)
```
## Look at assumptions of multicollinearity and influence

*Multicollinearity*
```{r}
#Variance Inflation Factors
vif(MULogit_multi)
```
#Cook's D plot
```{r}
plot(MULogit_multi, which=4, id.n=5, col="red") 
```

*Influence*
```{r}
#Cutoff: F(0.5, p, n-p), 4/n, 4/(n-p-1), n is the number of observations, p is the number of  coefficients
cutoff <- 0.0015 

#identify observations with a Cook's Distance greater than cutoff
obs_no <- as.data.frame(cooks.distance(MULogit_multi)) %>%
  mutate(obs_no=row_number()) %>%
  filter(`cooks.distance(MULogit_multi)` > cutoff)

#Let's exclude the values shown in the Cook's D plot, and see how the models compare.

MULogit_multi.modex <- update(MULogit_multi,subset=c(-obs_no$obs_no))
summary(MULogit_multi.modex)

#compare coefficients between models with and without influential observations, #caveat model number of observations is not the same
compareCoefs(MULogit_multi, MULogit) 
```


Did removing influential data affect the coefficients?

Answer: from the comparison results above, we can conclude that removing influential data largely affect the coefficients.

## Model fit of the full model
```{r}
#Various pseudo R squares, log likelihood, deviance, AIC, BIC
blr_model_fit_stats(MULogit_multi)

#Hosmer lemeshow goodness of fit test: a significant p value indicates a bad fit
blr_test_hosmer_lemeshow(MULogit_multi)
```



#testing effect modification
```{r}
Slp_PA_OBEM<- glm(sleep_deprive ~ PA+obese_new+(PA*obese_new)+overweight_new+(PA*overweight_new), data=yrbs_omit, family="binomial")
Slp_PA_OBLogit<-glm(sleep_deprive ~ PA+obese_new+overweight_new, data=yrbs_omit, family="binomial")
#likilihood ratio test: compare two  models
lrtest(Slp_PA_OBEM, Slp_PA_OBLogit)

Slp_PA_ObeEM<- glm(sleep_deprive ~ PA+obese_new+(PA*obese_new), data=yrbs_omit, family="binomial")
Slp_PA_ObeLogit<-glm(sleep_deprive ~ PA+obese_new, data=yrbs_omit, family="binomial")
#likilihood ratio test: compare two  models
lrtest(Slp_PA_OBEM, Slp_PA_ObeLogit)

Slp_PA_OEM<- glm(sleep_deprive ~ PA+overweight_new+(PA*overweight_new), data=yrbs_omit, family="binomial")
Slp_PA_OLogit<-glm(sleep_deprive ~ PA+overweight_new, data=yrbs_omit, family="binomial")
#likilihood ratio test: compare two  models
lrtest(Slp_PA_OEM, Slp_PA_OLogit)
```
The LR testing results shows that obese or overweight are not significant. Thus, we can exlude Effect Modification of obese/overweight out of our analysis.